Modeling Human Interactions in Collaborative Interactive Evolutionary Computation

### 2017-02-02 2017-GECCO-Modeling-User-Interactions-in-IEC/paper-body.tex
Starting Case Study
@JJ please have a look, I think only the EvoDrawings case study fits.
Done for today.

\section{Human-Centered IEC Framework}
% Comment: 
% Maybe in this case we should use the term software framework or architecture. 
The general goal of this research is to develop a human-centered \cite{gasson2003human} 
software framework for interactive evolutionary computation (IEC). 
The framework is a reusable architectural design together with an implementation \cite{campbell1991choices}, 
providing generalized components to developers of IEC systems. In the description we deliberately use the more general 
term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
could just be interacting naturally with their environment and with out knowing they are implicitly
evaluating a possible solution. The framework includes components that can be refined to increase
participation and also to minimize the amount of evaluations needed for the evolutionary process in a given IEC application. Software frameworks often have a vison or a way of thinking \cite{carneiro2010introducing} guiding their
design. Before entering to details the main design decisions of the framework are explained next:

@@

\item {\bf Users are volunteers} Users donate their computing resources, so they are unaccountable, 
  some times they try to game the system. Providers must actively promote and design the interactive
  system to engage volunteers \cite{oh2015clicking}.

@@

\item {\bf Context of use matters}
  \textit{Context} In computer science Fischer \cite{fischer2012context}
  defines context as \textit{``the interaction between humans and
  computers in socio-technical systems that takes place in a certain
  context referring to the physical and social situation in which
  computational devices and environments are embedded"}. 
  Fischer also identifies the important aspects to consider when the context is used: how it is
  the contextual data obtained, how the context is represented and what
  goals and purposes the context has when is used in a particular
  application.   An IEC system will  be used within a certain range of technical, physical and social or
  organizational environments \cite{maguire2001context} that may affect its use.
  The concept has been formally defined
  by ISO standard 9241-11 \cite{international1998iso} as \textit{``users,
  tasks and equipment (hardware, software and materials), and the
  physical and social environments in which a product is used"}.

@@

The human centered IEC architecture consists of three high level components depicted
in figure \ref{fig:uc_framework}:

@@

\section{Data Model Implementation}
The core of the framework is the data model because both the IEC system and the EA will 
depend on the application. In this section implementation details of the Data Model are presented. 
  \subsection{EvoSpace}
  \subsection{Graph}
  \subsection{PostgreSQL}

  \subsection{Interfaces}

@@

\section{Data Model Implementation}

\section{Case Studies}
\label{sec:evospace-i}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/interface.png}
    \caption{User interface of the EvoDrawings application.}
    \label{fig:web}
\end{figure}

@@

\subsection{Fitness Assignment}
\label{sec:assignment}
The ES-I platform employs a collaborative technique,
where several registered users assign a quality assessment to a single
phenotypes and then an aggregated fitness value is calculated. The fitness
assigned to each phenotypes depends on the taste of each particular user, 
resulting in a many-to-many relationship between users and phenotypes. 
Many systems query this user-phenotype relation to extract relevant
knowledge about the process and the population, for instance showing the
most popular phenotype, or the the user with more participation 
\cite{picbreeder}.
In order to do this, data about phenotypes 
must be permanently stored, even
if they are no longer in the active population.

\subsection{Collaboration}
\label{sec:col}
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those phenotypes they like, or by taking phenotypes
from their friend's collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, a phenotype sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a quality assessment to the phenotype using
a five star rating system or,
additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special folder that stores those phenotypes a user likes and wishes
to save. After the user finishes interacting with the phenotype
on the Wall, he can choose to retrieve a new one from the population.
At the left hand side, the web page shows the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, users can browse the content of each collection and from
there share images through their social network.
When a user hovers the cursor over a phenotypes a pop-up pane shows how many users have
liked the phenotype. The pane also includes a link to the phenotype's 
details, the parents, genetic operators that created it, and genealogy
information. This makes the assignment of fitness through the rating
system a {\em social} activity, pursuing the objective of this work,
which is to increase engagement of the users.
\subsubsection{Graph Model} 
\subsubsection{Human Interaction Model}

>>>>

\section{Human-Centered IEC Framework}
% Comment: 
% Maybe in this case we should use the term software framework or architecture. 
The general goal of this line of research is to develop a human-centered \cite{gasson2003human} 
software framework for interactive evolutionary computation (IEC). 
A framework is defined as a reusable architectural design together
with an implementation \cite{campbell1991choices}, in this case 
providing generalized components to developers of IEC systems. In the
description %below? If that's te chase, we _will_ - JJ
 we deliberately use the more general 
term Human instead of User, because humans are not always playing the role of users in software engineering sense, they
could just be interacting naturally with their environment and
unwittingly and at the same time they are implicitly
evaluating a possible solution.

%Which framework? Are we talking about a particular framework or that
%is the general rule for frameworks? - JJ
The framework includes components that can be refined to increase
participation and also to minimize the amount of evaluations needed for the evolutionary process in a given IEC application. Software frameworks often have a vison or a way of thinking \cite{carneiro2010introducing} guiding their
design. Before diving into details the main design decisions of the
framework are explained next:  % constraints? Rules? You don't decide
                               % that users are humans, they simply
                               % are - JJ

@@

\item {\bf Users are volunteers} Users donate their computing resources, so they are unaccountable, 
  some times they try to game the system. Providers must actively promote and design the interactive
  system to engage volunteers \cite{oh2015clicking}. % Define
                                % providers somewhere. Experiment
                                % owners? - JJ

@@

\item {\bf Context of use matters}
  \textit{Context} In computer science Fischer \cite{fischer2012context}
  defines context as \textit{``the interaction between humans and
  computers in socio-technical systems that takes place in a certain
  context referring to the physical and social situation in which
  computational devices and environments are embedded''}. 
  Fischer also identifies the important aspects to consider when the context is used: how it is
  the contextual data obtained, how the context is represented and what
  goals and purposes the context has when is used in a particular
  application.   An IEC system will  be used within a certain range of technical, physical and social or
  organizational environments \cite{maguire2001context} that may affect its use.
  The concept has been formally defined
  by ISO standard 9241-11 \cite{international1998iso} as \textit{``users,
  tasks and equipment (hardware, software and materials), and the
  physical and social environments in which a product is used"}.

@@

% Say something about how this is important for the framework and how
% you incorporate it - JJ
The human centered IEC architecture consists of three high level components depicted
in figure \ref{fig:uc_framework}:

@@

\section{Data Model Implementation}
The core of the framework is the data model because both the IEC system and the EA will 
depend on the application. In this section implementation details of the Data Model are presented.
Three database models are employed to store different elements of the system:

  \subsection{EvoSpace-Redis}
EvoSpace is a population store \cite{Evospace}  for the development of 
evolutionary algorithms that are intended to run on a cloud computing model. 
The population is decoupled from any particular evolutionary algorithm. 
Candidate solutions are stored as of objects in a population, and they can be withdrawn, 
processed and replaced using a specified set of methods \cite{GValdez2015}. The population
is stored in-memory, using the Redis key-value database. Redis was chosen over a relational 
database management system, because it provides a hash based implementation of sets and 
queues which are natural data structures for the EvoSpace model. Basicaly a sample of 
candidate solutions are retieved from the server, evaluated, and then sent back. 
The same operations are used to Evolve the population. In EvoSpace individuals replaced 
in the population are stored indefinitly, to permit users to store permalinks to them.
Implementation details are presented in \cite{garcia2013evospace}.

\subsection{Graph}
  In order to deal with large datasets of connected data found in these systems,
  graph databases \cite{angles2012comparison} have been proposed as an alternative to relational databases 
  which have performance limitations when dealing with highly connected data \cite{holzschuher2013performance}.
  A graph is proposed for modeling the social network of users, their interaction with 
  candidate solutions, and the relationships between them in the population.
  The graph database system used in the implementation is Neo4J as it is
  a scalable solution \cite{miller2013graph,holzschuher2013performance}, and well supported and documented
  in PaaS infrastructures like Heroku. The Cypher query language it used to retrieve views from the graph.
  An exaplmple query is shown in figure \ref{fig:cypher} where the relations between users and solutions 
  are presented.

  \begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/gui-neo.png}
    \caption{Example query in Cypher.}
    \label{fig:cypher}
  \end{figure}

\subsection{PostgreSQL}
  The PostgreSQL relational database system, is also employed because user sessions and authentication,
  as well as dynamic web pages are handled directly by the Django web framework \cite{garcia2013evospace}.

\subsection{Interfaces}
%They are Sensor and Web,

@@

\section{Case Studies}
\label{sec:evospace-i}

\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/interface.png}
    \caption{User interface of the EvoDrawings application.}
    \label{fig:web}
\end{figure}

@@

\subsection{Fitness Assignment}
\label{sec:assignment}
The ES-I % You should use always the same acronym, not start using it
         % here - JJ
 platform employs a collaborative methodology for performing fitness
 assignment. Several registered users assign a quality assessment to a single
phenotype and then an aggregated fitness value is calculated,
depending on the tastes of each particular user,  % how do you define
                                % "taste"? - JJ
resulting in a many-to-many relationship between users and
phenotypes. %scored by them? - JJ 
Many systems % formerly published methods? Other researchers? - JJ
query this user-phenotype relation to extract relevant
knowledge about the process and the population, for instance showing the
most popular phenotype, or the the user with more participation 
\cite{picbreeder}.
In order to do this, data about phenotypes % metadata better? - JJ
must be permanently stored, even
if they are no longer in the active population.

\subsection{Collaboration}
\label{sec:col}
Users need to authenticate themselves to the system using their
Facebook account. In that sense, some users might not be interested
either because they do not want to give that that or simply because
they do not use that particular social network. Even as we might lose
some users that way, the additional information we obtain for scoring
phenotypes more than compensates for that. % Added this, check it
                                % out. 
After entering the web application by using their Facebook account,
users can collaborate with their Facebook friends, 
sharing those phenotypes they like, or by taking phenotypes
from their friend's collections by using the web interface depicted 
in Figure \ref{fig:web}.
At the top left corner a list of Facebook friends is presented
to encourage users to interact with the system. In the central 
\emph{ Wall } area, a phenotype sampled from the population that is
being evolved via the evolutionary algorithm 
is shown to the user.
Here, the user can interact with the system in two ways.
First, he can assign a quality assessment % Why don't you use other
                                % thing instead of quality? We are not
                                % talking about quality, but about
                                % tastes or similar - JJ
 to the phenotype using
a five star rating system or,
additionally, a user can choose to add an image to one of their \emph{Collections}.
A collection is a special folder that stores those phenotypes a user likes and wishes
to save. After the user finishes interacting with the phenotype
on the Wall, he can choose to retrieve a new one from the population.
At the left hand side of \ref{fig:web}, the web page shows the \emph{Collections} section.
The user can create several collections, to group and organize his favorite 
artifacts. Moreover, users can browse the content of each collection and from
there share images through their social network.
When a user hovers the cursor over a phenotypes a pop-up pane shows how many users have
liked the phenotype. The pane also includes a link to the phenotype's 
details, the parents, genetic operators that created it, and genealogy
information. This makes the assignment of fitness through the rating
system a {\em social} activity, pursuing the objective of this work,
which is to increase engagement of the users.
\subsubsection{Graph Model}

\subsubsection{Gammification}

\subsubsection{Setup}
\subsubsection{Results}

\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/weibull_1.png}
    \caption{Experiment 1. Weibull Fit}
    \label{fig:}
\end{figure}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/weibull_2.png}
    \caption{Experiment 2. Weibull Fit}
    \label{fig:}
\end{figure}
\begin{figure}[!t]
    \centering
        \includegraphics[width=3.5in]{img/weibull_3.png}
    \caption{Experiment 3. Weibull Fit}
    \label{fig:}
\end{figure}
